---
title: "p8105_HW1_mk4208"
author: "Meeraj Kothari"
date: 2019-09-18
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading packages

```{r}
library(tidyverse)
```


# Problem 1 

## Creating a dataframe and taking the means of the variables in the data frame

```{r}
# Creating a data frame

set.seed(100)
df = tibble(
  random = rnorm(8),
  vec_logical = random > 0,
  vec_char = c("a", "b", "c", "e", "f", "g", "h", "i"),
  vec_factor = factor(c("M", "F", "X", "M", "F", "X", "M", "F"))
)

# Taking the mean of each variable in the dataframe 

mean(pull(df, random)) # WORKS 

mean(pull(df, vec_logical)) # WORKS 

mean(pull(df, vec_char)) # DOES NOT WORK 

mean(pull(df, vec_factor)) # DOES NOT WORK
```

## Coercing the logical, character and factor variables from the dataframe to numeric and then taking the means 

###### Writing a code chunk that applies as.numeric function to the logical, character, and factor variables 
```{r}

mean(as.numeric(pull(df, vec_logical))) 
```
as.numeric converts FALSE to 0 and TRUE to 1 and the mean function takes the mean of these newly converted values 

```{r}
mean(as.numeric(pull(df, vec_char)))
```
as.numeric is unable to convert the character strings to numeric values and hence the mean function returns NA

```{r}
mean(as.numeric(pull(df, vec_factor))) 
```
as.numeric converts the first level of the factor to 1, second level of the factor to 2 and third level of the factor to 3 and the mean function takes the mean of these newly converted values

## The next code chunk does the following: 
###### 1. Converts the logical vector to numeric, and multiplies the random sample by the result
###### 2. Converts the logical vector to a factor, and multiplies the random sample by the result
###### 3. Converts the logical vector to a factor and then converts the result to numeric, and multiplies the random sample by the result

```{r}
# 1. Converting the logical vector to numeric and multiplying the random sample by the result 

as.numeric(pull(df, vec_logical)) * pull(df, random) 

# 2. Converting the logical vector to a factor and multiplying the random sample by the result 

as.factor(pull(df, vec_logical)) * pull(df, random)

# 3. Converting the logical vector to a factor and then converting the result to numberic and multiplying the random sample by the result 

as.numeric(as.factor(pull(df, vec_logical))) * pull(df, random)

```

# Problem 2 

Creating a data frame comprised of:
  x: a random sample of size 500 from a standard Normal distribution.
  y: a random sample of size 500 from a standard Normal distribution.
  A logical vector indicating whether x + y > 1.
  A numeric vector created by coercing the above logical vector.
  A factor vector created by coercing the above logical vector.

```{r}
# Creating a data frame 
df2 = tibble(
  x = rnorm(500),
  y = rnorm(500),
  vec_logical = x + y > 1,
  vec_numeric = as.numeric(vec_logical),
  vec_factor = as.factor(vec_logical)
)
```

Here is a short description of the dataframe created above: 
This dataframe has `r nrow(df2)` rows and `r ncol(df2)` columns. 
The mean, median and standard deviation of 'x' variable is `r mean(pull(df2, x))`, `r median(pull(df2, x))`, and `r sd(pull(df2, x))` respectively. The proportion of cases with x + y > 1 is `r mean(pull(df2, x))` 

```{r}
# Creating a scatter plot of  y vs x; color points using the logical variable
plot1 = ggplot(df2, aes(x=x, y=y, color=vec_logical)) + 
  geom_point() 

print(plot1)
```
The color scale is categorical and the color assigned to FALSE values is orange and TRUE values is green

```{r}
# Creating a scatter plot of  y vs x; color points using the numeric variable
plot2 = ggplot(df2, aes(x=x, y=y, color=vec_numeric)) + 
  geom_point() 

print(plot2) 
```
The color scale is continuous and the colors are assigned from low (0.00) = BLACK to high (1.00) = BLUE 

```{r}
# Creating a scatter plot of  y vs x; color points using the factor variable
plot3 = ggplot(df2, aes(x=x, y=y, color=vec_factor)) + 
  geom_point()

print(plot3) 
```
The color scale is categorical and the color assigned to FALSE values is orange and TRUE values is green

```{r}
# Exporting the first scatterplot 
ggsave("plot1.png", plot=plot1)
```



